Ethereum: A Secure Decentralised Transaction Progression Platform

### 2014-04-03 Paper.tex
Execution model.

\title{Ethereum: A Secure Decentralised Compute Platform}
\author{
    Gavin Wood\\
    CTO, Ethereum Project\\
    gavin@ethereum.org
}
\begin{document}

@@

In addition to the system state $\mathbb{S}$, there are several pieces of important, immutable, information used in the execution environment that the execution agent must provide:

\begin{itemize}
\item $a$, the address of the account which own the code that is executing.
\item $o$, the sender address of the transaction that originated this execution.
\item $p$, the price of gas in the transaction that originated this execution.
\item $D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
\item $s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
\item $v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
\item $H$, the block header of the previous block.
\item $H'$, the block header of the current block.
\item $n$, the block number of the current block.
\end{itemize}

The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$, given the required information:

\begin{equation}
\mathbb{S}' \equiv X(\mathbb{S}, a, o, p, D, s, v, H, H', n)
\end{equation}

@@

\begin{enumerate}
\item Let $\mathbb{S}' := \mathbb{S}$, the current state.
\item Let $PC := 0$, the program counter.
Let it be known that all storage operations operate on TO's state.
Note that the memory is empty, thus (last used index + 1) = 0.
\item Set TXDATA to the first INSIZE bytes of memory starting from INOFFSET in the caller memory.
\item Repeat
\begin{itemize}
\item Calculate the cost of the current instruction (see below), set to $C$.
\item If the instruction is invalid or STOP, goto step 6.
\item If $GAS < C$ then $GAS := 0$; $S := ROLLBACK$ and evaluation finishes, returning false. 
\item $GAS := GAS - C$
\item Apply the instruction.
\end{itemize}
Until an operation execution error has occured or the instruction is STOP, RETURN or SUICIDE.
\item If the output data of the call (R) is specified (through RETURN), then let the OUTSIZE bytes of caller memory beginning at OUTOFFSET.
\item Returns true.
\end{enumerate}

When a contract address receives a transaction, a virtual machine is initiated with the contract's state.

@@

There exists a stack of variable size that stores 256-bit values at each location. (Note: most instructions operate solely on the stack altering its state in some way.)

T'[i] is the ith item counting down from the top of the pre-stack (i.e. the stack immediately after instruction execution), with the top of the stack corresponding to i == 0.

T[i] is the ith item counting down from the top of the post-stack (i.e. the stack immediately prior to instruction execution), with the top of the stack corresponding to i == 0.

The exists a permanent store addressable by a 256-bit value that stores 256-bit values at each location.

S'[i] is the permanent store (sometimes refered to as 'state' or 'store') of the VM at index i counting from zero PRIOR to instruction execution.

S[i] is the permanent store (sometimes refered to as 'state' or 'store') of the VM at index i counting from zero AFTER to instruction execution.

The exists a temporary store addressable by a 256-bit value that stores 256-bit values at each location.

M'[i] is the temporary store (sometimes refered to as 'memory') of the VM at index i counting from zero PRIOR to instruction execution.

M[i] is the temporary store (sometimes refered to as 'memory') of the VM at index i counting from zero AFTER instruction execution.

C is the code, a byte sequence.

PC' is the program counter PRIOR to instruction execution.

PC is the program counter AFTER instruction execution.

FEE(I, S', P', D) is the fee associated with the execution of instruction I with a machine of stack S', permanent store P' and which has already completed D operations.

It is defined as F where:

TODO: No it's not!

IF I == SSTORE AND P[ S'[0] ] != 0 AND S'[1] == 0 THEN
    F = S + dataFee - memoryFee
IF I == SSTORE AND P[ S'[0] ] == 0 AND S'[1] != 0 THEN
    F = S + dataFee + memoryFee
IF I == SLOAD
    F = S + dataFee
IF I == EXTRO OR I == BALANCE
    F = S + extroFee
IF I == MKTX
    F = S + txFee
IF I == SHA256 OR I == SHA3 OR I == RIPEMD160 OR I == ECMUL OR I == ECADD
    OR I == ECSIGN OR I == ECRECOVER OR I == ECVALID THEN
    F = S + cryptoFee

B[ A ] is the balance of the address given by A, with A interpreted as an address.

ADDRESS is the address of the present contract.

10. b. Initial Operation

STEPSDONE := 0
PC' := 0
FOR ALL i: M' is initialised to the empty byte sequence.
T' is initialised such that its cardinality is zero (i.e. the stack starts empty).
S' must be equal to the value of P when the previous execution halted.

10. c. General Operation

The given steps are looped:
1. Execution halts if B[ ADDRESS ] < F( C[PC'], S', P', STEPSDONE )
2. B[ ADDRESS ] := B[ ADDRESS ] - F( P'[PC'], S', P', STEPSDONE )
3. The operation given by P'[PC'] determines PC, P, T, S.
4. PC' := PC; P' := P; T' := T; S' := S; STEPSDONE := STEPSDONE + 1

10. d. VM Operations

@@

If PC is not defined explicitly, then it must be assumed PC := PC' + 1. Exceptions are PUSH, JUMP and JUMPI.

>>>>

\title{Ethereum: A Secure Decentralised Transaction Progression Platform}
\author{
    Gavin Wood\\
    CTO, Ethereum Project\\
    gavin@ethereum.org
}
\begin{document}

@@

In addition to the system state $\mathbb{S}$, and the remaining gas for computation $g$, there are several pieces of important information used in the execution environment that the execution agent must provide; these are contained in the tuple $I$:

\begin{itemize}
\item $I_a$, the address of the account which owns the code that is executing.
\item $I_o$, the sender address of the transaction that originated this execution.
\item $I_p$, the price of gas in the transaction that originated this execution.
\item $I_D$, the byte array that is the input data to this execution; if the execution agent is a transaction, this would be the transaction data.
\item $I_s$, the address of the account which caused the code to be executing; if the execution agent is a transaction, this would be the transaction sender.
\item $I_v$, the value passed to this account as part of the same procedure as execution; if the execution agent is a transaction, this would be the transaction value.
\item $I_H$, the block header of the previous block.
\item $I_{H'}$, the block header of the current block.
\item $I_n$, the block number of the current block.
\item $I_C$, the byte array that is the machine code to be executed.
\end{itemize}

The execution model defines the function $X$, which can compute the resultant state $\mathbb{S}'$ and the remaining gas $g'$, given these definitions:

\begin{equation}
(\mathbb{S}', g') \equiv \Xi(\mathbb{S}, g, I)
\end{equation}

@@

We must now define the $\Xi$ function. In most practical implementations this will be modelled as an iterative progression of the pair comprising the full system state, $\mathbb{S}$ and the machine state, $\mathbb{M}$. Formally, we define it recursively with a function $X$. This uses an iterator function $O$ (which defines the result of a single cycle of the state machine) together with functions $Z$ which determines if the present state is an exceptional halting state of the machine and $H$, specifying if the present state is a normal halting state of the machine:

\begin{eqnarray}
\Xi(\mathbb{S}, g, I) & \equiv & X(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}_g & \equiv & g \\
\mathbb{M}_{pc} & \equiv & 0 \\
\mathbb{M}_M & \equiv & [0, 0, ...] \\
\mathbb{M}_i & \equiv & -1 \\
\mathbb{M}_T & \equiv & [] \\
X(\mathbb{S}, \mathbb{M}, I) & \equiv & \begin{cases}
(\mathbb{S}, \mathbb{M}, I) & \text{if} \quad Z(\mathbb{S}, \mathbb{M}, I)\\
O(\mathbb{S}, \mathbb{M}, I) & \text{if} \quad H(\mathbb{S}, \mathbb{M}, I)\\
X(O(\mathbb{S}, \mathbb{M}, I)) & \text{otherwise}\\
\end{cases}
\end{eqnarray}

The machine state $\mathbb{M}$ is defined as the tuple $(g, pc, M, i, T)$ which are the gas available, the program counter, the memory contents, the highest used memory index, and the stack contents. The memory contents $\mathbb{M}_M$ are a series of zeroes of size $2^{256}$.

For the ease of reading, the instruction mnemonics, written in smallcaps (\eg {\small ADD}), should be interpreted as their numeric equivalents; the full table of instructions and their specifics is given in Appendix \ref{app:vm}.

\subsubsection{Exceptional Halting}

For the purposes of defining $H$ and $O$, we define $o$ as the current operation to be executed:

\begin{equation}
o \equiv \begin{cases} I_C[\mathbb{M}_{pc}] & \text{if} \quad \mathbb{M}_{pc} < \lVert I_C \rVert \\
\text{\small STOP} & \text{otherwise}
\end{cases}
\end{equation}

We also assume the fixed amounts of $\mathbf{\delta}$ and $\mathbf{\alpha}$, specifying the stack movement items removed and added, both subscriptable on the instruction and an instruction cost function $C$ evaluating to the full cost, in gas, of executing the given instruction. We also institute a function

For calculating the exceptional halting function $Z$, we assume a secondary, instruction-specific halting function $Y$, which identifies edge cases specific to particular instructions (such as the division by zero or initiating a message call with insufficient funds). Thus $Z$ is defined as:

\begin{equation}
\begin{array} {lll}
Z(\mathbb{S}, \mathbb{M}, I) & \equiv & Y(\mathbb{S}, \mathbb{M}, I) \\
& & \vee \quad \mathbb{M}_g < C(\mathbb{S}, \mathbb{M}, I) \\
& & \vee \quad \lVert\mathbb{M}_T\rVert < \mathbf{\delta}_o
\end{array}
\end{equation}

This states that the execution is in an exceptional halting state if there is an instruction-specific issue, if there is insufficient gas or if there are insufficient stack items.

\subsubsection{The Execution Cycle}

Stack items are added or removed from the left-most, lower-indexed portion of the series; all other items remain unchanged:

\begin{eqnarray}
O(\mathbb{S}, \mathbb{M}, I) & \equiv & (\mathbb{S}', \mathbb{M}', I) \\
\Delta & \equiv & \mathbf{\alpha}_o - \mathbf{\delta}_o \\
\lVert\mathbb{M}'_{T}\rVert & \equiv & \lVert\mathbb{M}_{T}\rVert + \Delta \\
\quad \forall i \in [\mathbf{\alpha}_o, \lVert\mathbb{M}'_T\rVert): \mathbb{M}'_{T_i} & \equiv & \mathbb{M}_{T_{i+\Delta}}
\end{eqnarray}

The gas is reduced by the instruction's gas cost and for most instructions, the program counter increments on each cycle, for the three exceptions, we assume a function $J$ which evaluates to the according value:

\begin{eqnarray}
\mathbb{M}'_{g} \equiv \mathbb{M}_{g} - C(\mathbb{S}, \mathbb{M}, I) \\
\mathbb{M}'_{pc} \equiv \begin{cases}
J(\mathbb{M}) & \text{if} \quad o \in \{ \text{\small JUMP, JUMPI} \} \\
\mathbb{M}_{pc} + p & \text{if} \quad o \in [\text{\small PUSH1}, \text{\small PUSH32}] \\
\mathbb{M}_{pc} + 1 & \text{otherwise}
\end{cases}
\end{eqnarray}

where $p$ is the byte size of the push instruction, defined as:

\begin{equation}
p \equiv o - \text{\small PUSH1} + 2
\end{equation}

In general, we assume the memory and system state don't change:

\begin{eqnarray}
\mathbb{M}'_M & \equiv & \mathbb{M}_M \\
\mathbb{M}'_i & \equiv & \mathbb{M}_i \\
\mathbb{S}' & \equiv & \mathbb{S}
\end{eqnarray}

Certain instructions do; alter them---such changes are noted in Appendix \ref{app:vm}.

@@

@@

If PC is not defined explicitly, then it must be assumed PC := PC' + 1. Exceptions are .
